Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-04 paper/paper.tex
Intro again

Our work represents a substantial refinement of {\sc brahma}~\cite{brahma},
which is also the closest work to ours is.  {\sc brahma} is also a method
for synthesizing loop-free bitvector programs.  In~\cite{brahma}, the
authors characterise the program synthesis problem as a second-order
$\exists \forall$ formula, and introduce the CEGIS algorithm for solving
these formulae.  This algorithm reduces the satisfiability of the $\exists
\forall$ formula to the satisfiability of several first-order formulae, each
of which can be decided using an off-the-shelf SMT solver.  {\sc Brahma}
takes as input a specifcation (given as an SMT formula) and a library of
components (one component is more or less a single machine instruction).  It
then finds a permutation of the component library that satisfies the
specification.  This permutation is the final synthesised program.  In the
event that the library does not contain enough components to synthesise a
correct program, the user is told that the library is too small and is
required to add more components until the synthesis succeeds.  Each round of
synthesis amounts to solving an $\exists \forall$ formula that is quadratic
in the size of the component library.

@@

Another successful approach to program synthesis has been \emph{program
sketching}, in particular the {\sc sketch} tool~\cite{sketch}.  With this
method the user supplies a program sketch, written in a C-like language,
together with either a specification (written in the same language) or a
suite of unit tests.  The sketch is a skeleton program which contains much
of the high level structure of the final program, but with several ``holes''
that {\sc sketch} fills in to meet the specification, again using the CEGIS
algorithm.

Program synthesis is strongly related to superoptimisation~\cite{superoptimisation},
which is the problem of taking a segment of code, usually written in C, and finding
a minimal program which has the same functionality.  The {\sc aha}~\cite{aha} tool
and the GNU superoptimiser~\cite{gnu-superoptimiser} both do superoptimisation of
C programs, both by enumerating possible programs and testing them on suitable inputs.
This search can be compiled down to native code using an optimising compiler,
so for short programs the method is extremely fast, however because of the exponential
blowup of the program space it is impractical for synthesising longer programs.
We incorporate this technique by using explicit state model checking in parallel
with symbolic model checking.

\section{The Basic Synthesis Algorithm}

@@

\bigskip

\begin{tabular}{l|l|l|l|l}
Problem & {\sc Brahma} & {\sc Kalashnikov} & {\sc Kalashnikov} & {\sc Kalashnikov} \\
        &               &                    & (no width)         & (CBMC only) \\
\hline
p1 & 3.20s &{\bf 1.83s} &23.54s &23.54s \\
p2 & 3.60s &{\bf 1.49s} &1271.12s &1271.12s \\
p3 & {\bf 1.40s} &1.60s &4.09s &4.09s \\
p4 & 3.30s &{\bf 1.86s} &79.72s &79.72s \\
p5 & 2.20s &{\bf 1.98s} &12.16s &12.16s \\
p6 & 2.40s &{\bf 1.51s} &104.19s &104.19s \\
p7 & {\bf 1.00s} &4.04s &- &- \\
p8 & {\bf 1.40s} &4.31s &- &- \\
p9 & {\bf 5.80s} &35.09s &505.10s &505.10s \\
p10 & 76.10s &{\bf 9.80s} &- &- \\
p11 & 57.10s &{\bf 9.40s} &{\bf 9.40s} &{\bf 9.40s} \\
p12 & 67.80s &{\bf 21.43s} &{\bf 21.43s} &{\bf 21.43s} \\
p13 & 59.60s &{\bf 27.74s} &{\bf 27.74s} &{\bf 27.74s} \\
p14 & 118.90s &{\bf 54.71s} &{\bf 54.71s} &{\bf 54.71s} \\
p15 & 62.30s &{\bf 43.49s} &{\bf 43.49s} &{\bf 43.49s} \\
p16 & 78.10s &{\bf 27.41s} &{\bf 27.41s} &{\bf 27.41s} \\
p17 & 45.90s &{\bf 27.73s} &{\bf 27.73s} &{\bf 27.73s} \\
p18 & 34.70s &{\bf 5.55s} &{\bf 5.55s} &{\bf 5.55s} \\
p19 & {\bf 108.40s} &- &- &- \\
p20 & {\bf 28.30s} &- &- &- \\
p21 & {\bf 279.00s} &- &- &- \\
p22 & {\bf 1668.00s} &- &- &- \\
p23 & {\bf 224.90s} &- &- &- \\
p24 & {\bf 2778.70s} &- &- &- \\
\end{tabular}

>>>>

Our work represents a substantial refinement of {\sc brahma}~\cite{brahma},
which is also the closest work to ours.  {\sc Brahma} also focuses on
synthesizing loop-free machine-integer programs.  In~\cite{brahma}, the
authors characterise the program synthesis problem as a second-order
$\exists \forall$ formula, and introduce the CEGIS algorithm for solving
these formulae.  This algorithm reduces the satisfiability of the $\exists
\forall$ formula to the satisfiability of several first-order formulae, each
of which can be decided using an off-the-shelf SMT solver.  {\sc Brahma}
takes as input a specifcation (given as an SMT formula) and a library of
components (one component is more or less a single machine instruction).  It
then finds a permutation of the component library that satisfies the
specification.  This permutation is the final synthesised program.  In the
event that the library does not contain enough components to synthesise a
correct program, the user is told that the library is too small and is
required to add more components until the synthesis succeeds.  Each round of
synthesis amounts to solving an $\exists \forall$ formula that is quadratic
in the size of the component library.

@@

Another successful approach to program synthesis has been \emph{program
sketching}, in particular the {\sc sketch} tool~\cite{sketch}.  This method
requires the programmer to provide a \emph{program sketch}, written in a
C-like language, together with either a specification (written in the same
language) or a suite of unit tests.  The sketch is a skeleton program which
contains much of the high-level structure of the final program, but with
several ``holes'', which {\sc sketch} fills in to meet the specification,
again using the CEGIS algorithm.

Program synthesis is strongly related to superoptimisation~\cite{superoptimisation},
which is the problem of taking a segment of code, usually written in C, and finding
a minimal program which has the same functionality.  The {\sc aha}~\cite{aha} tool
and the GNU superoptimiser~\cite{gnu-superoptimiser} both perform superoptimisation of
C programs, both by enumerating possible programs and testing them on suitable inputs.
This search can be compiled down to native code using an optimising compiler,
so for short programs the method is extremely fast, however because of the exponential
blowup of the search space it is impractical for synthesising longer programs.
We incorporate this technique by using explicit-state model checking in parallel
with symbolic model checking.

%\paragraph{Outline} We give details of our new algorithm in
%Sec.~\ref{sec:algorithm}.

\section{The Basic Synthesis Algorithm}
\label{sec:algorithm}

@@

\begin{figure}
\begin{center}
\begin{tabular}{l|l|l|l|l}
Problem & {\sc Brahma} & {\sc Kalashnikov} & {\sc Kalashnikov} & {\sc Kalashnikov} \\
        &               &                    & (no width)         & (CBMC only) \\
\hline
p1 & 3.20s &{\bf 1.83s} &23.54s &8.07s \\
p2 & 3.60s &{\bf 1.49s} &1271.12s &8.19s \\
p3 & {\bf 1.40s} &1.60s &4.09s &5.90s \\
p4 & 3.30s &{\bf 1.86s} &79.72s &8.15s \\
p5 & 2.20s &{\bf 1.98s} &12.16s &8.10s \\
p6 & 2.40s &{\bf 1.51s} &104.19s &8.17s \\
p7 & {\bf 1.00s} &4.04s &- &14.49s \\
p8 & {\bf 1.40s} &4.31s &- &14.20s \\
p9 & {\bf 5.80s} &35.09s &505.10s &13.45s \\
p10 & 76.10s &{\bf 9.80s} &- &19.14s \\
p11 & 57.10s &{\bf 9.40s} &{\bf 9.40s} &32.98s \\
p12 & 67.80s &{\bf 21.43s} &{\bf 21.43s} &38.23s \\
p13 & 59.60s &27.74s &27.74s &{\bf 13.16s} \\
p14 & 118.90s &54.71s &54.71s &{\bf 52.69s} \\
p15 & 62.30s &{\bf 43.49s} &{\bf 43.49s} &44.31s \\
p16 & 78.10s &{\bf 27.41s} &{\bf 27.41s} &45.84s \\
p17 & 45.90s &{\bf 27.73s} &{\bf 27.73s} &46.63s \\
p18 & 34.70s &{\bf 5.55s} &{\bf 5.55s} &15.63s \\
p19 & {\bf 108.40s} &- &- &- \\
p20 & {\bf 28.30s} &- &- &- \\
p21 & {\bf 279.00s} &- &- &- \\
p22 & {\bf 1668.00s} &- &- &- \\
p23 & {\bf 224.90s} &- &- &- \\
p24 & {\bf 2778.70s} &- &- &- \\
\end{tabular}

\end{center}

\label{fig:results-table}
\caption{Times for synthesis of machine integer benchmarks from {\sc brahma} paper}
\end{figure}
